---
title: "Segmentation Example"
author: "Fareeza Khurshed"
date: "4/13/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=TRUE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(factoextra)
library(standardize)
library(fmsb)


#read in all raw data
rawData <- read_csv("https://raw.githubusercontent.com/statgeek/data/master/census%20FSA.csv")

#code for graph later - included her for simplicity
create_beautiful_radarchart <- function(data, color = "#00AFBB", 
                                        vlabels = colnames(data), vlcex = 0.7,
                                        caxislabels = NULL, title = NULL, ...){
  radarchart(
    data, axistype = 1,
    # Customize the polygon
    pcol = color, pfcol = scales::alpha(color, 0.5), plwd = 2, plty = 1,
    # Customize the grid
    cglcol = "grey", cglty = 1, cglwd = 0.8,
    # Customize the axis
    axislabcol = "grey", 
    # Variable labels
    vlcex = vlcex, vlabels = vlabels,
    caxislabels = caxislabels, title = title, ...
  )
}


```

## Segmentation Examples

Examine the data and then do some data cleaning. 

* Removes redundant variables from a contextually basis
* Filters data to Alberta to ensure relevance
* Ensure that no data is missing

This data is Census 2016 data


```{r summary}


#remove columns that will not be used for modeling
dataP1 <- rawData %>%
  select(-c(year, geo_code, geo_level, GNR, GNR_LF, data_quality_flag, alt_geo_code,
            pct_age_0_14, pct_unemployed, pct_cdn_citizens, pct_immigrants, avg_census_family_size,
            median_income_recipients, pct_participation_rate, pct_home_owners, pct_education_none,
            num_dwellings_occupied ), 
         -starts_with("pct_age_"),
         -starts_with("pct_LIM_AT_")
  ) %>%
  filter(substr(FSA, 1, 1) == "T") %>%
  column_to_rownames(var="FSA")

#remove all columns where missing values and convert to data frame
dataP2 <- dataP1 %>% filter(complete.cases(.))

summary(dataP2)
```

## K Means Model

Trying it out without further processing:

```{r attempt1, echo=FALSE}
demo <- kmeans(dataP2, 3)

fviz_cluster(demo, data = dataP2,
             palette = c("#00AFBB","#2E9FDF", "#E7B800"),
             ggtheme = theme_minimal(),
             main = "Partitioning Clustering Plot"
)

```

## Standardization!

* Standardize variables and repeat

```{r}
dataP3 <- as_tibble(scale(dataP2))

demo2 <- kmeans(dataP3, 3)

fviz_cluster(demo2, data = dataP3,
             palette = c("#00AFBB","#2E9FDF", "#E7B800"),
             ggtheme = theme_minimal(),
             main = "Partitioning Clustering Plot"
)
```
## Hierarchical Modeling

```{r}
# Compute hierarchical clustering and cut into 3 clusters
ha1_model <- hcut(dataP3, k = 3, stand = TRUE)

# Visualize
fviz_dend(ha1_model, rect = TRUE, cex = 0.5,
          k_colors = c("#00AFBB","#2E9FDF", "#E7B800"))
```
## Results comparison

```{r}
#add cluster to data
dataP4 <- dataP3
dataP4$cluster_kmeans <- demo2$cluster
dataP4$cluster_hier <- ha1_model$cluster

table(dataP4$cluster_kmeans, dataP4$cluster_hier)
```

## Optimal Number of clusters?

```{r}
#optimal number of clusters

fviz_nbclust(dataP3, kmeans, method = "silhouette")
```
## Now what?

* Visual your data

```{r}

#source: https://www.datanovia.com/en/blog/beautiful-radar-chart-in-r-using-fmsb-and-ggplot-packages/

#visualize using a radar chart
df_scaled <- dataP3

# Variables summary
# Get the minimum and the max of every column  
col_max <- apply(df_scaled, 2, max)
col_min <- apply(df_scaled, 2, min)
# Calculate the average profile 
col_mean <- apply(df_scaled, 2, mean)

#cluster1

summary<-dataP4 %>% 
  group_by(cluster_kmeans) %>%
  summarise(across(pop_2016:pct_non_movers, ~ mean(.x, na.rm = TRUE))) %>%
  select(-cluster_kmeans)

# Put together the summary of columns
col_summary <- t(data.frame(Max = col_max, Min = col_min, Average = col_mean))


# Bind variables summary to the data
df_scaled2 <- as.data.frame(rbind(col_summary, summary))


# Define colors and titles
colors <- c("#00AFBB", "#E7B800", "#FC4E07")
titles <- c("Cluster 1", "Cluster 2", "Cluster 3")

# Reduce plot margin using par()
# Split the screen in 3 parts
op <- par(mar = c(1, 1, 1, 1))
par(mfrow = c(1,3))

# Create the radar chart
for(i in 1:3){
  create_beautiful_radarchart(
    data = df_scaled2[c(1, 2, i+2), ],
    color = colors[i], title = titles[i]
  )
}
par(op)

```

